---
title: "Statistical Inference Course Project"
author: "Daniel Resende"
date: "February 20, 2015"
output: html_document
---

```{r libraries, echo=FALSE}
library(ggplot2)
```

### Overview
In this report I have made some simulations and compared them to some theoretical values concerning sample variances, sample means and its distribution. It works.

```{r simulation}
lambda <- .2
nosim <- 1000
n <- 40
data <- rexp(nosim * n, lambda)
m <- matrix(data, nosim)
```
I have made `r nosim` simulations with size of `r n` each. The values were generated using rexp distribution with lambda = `r lambda`. They were put in a matrix m.

### 1. Sample x Theoretical means
```{r sampleMeans, fig.width=16}
sampleMeans <- apply(m, 1, mean)
g <- ggplot(data.frame(sampleMeans), aes(sampleMeans)) +
    geom_histogram(aes(y=..density..), binwidth=.1) +
    geom_density() +
    geom_vline(aes(xintercept=1/lambda, colour="Theoretical", linetype="Theoretical"), show_guide = TRUE) +
    geom_vline(aes(xintercept=mean(data), colour="Simulation", linetype="Simulation"), show_guide = TRUE) +
    scale_colour_manual(name="Means", values=c(Theoretical="blue", Simulation="red")) +
    scale_linetype_manual(name="Means", values=c(Theoretical="dashed", Simulation="dashed"), guide=FALSE) +
    ggtitle("Sample x Theoretical means") +
    xlab("Mean") +
    ylab("Density")
print(g)
```
The theoretical mean is 1/lambda = `r 1/lambda` (blue dasehd line). In this simulation, our mean of sample means was `r mean(sampleMeans)` (red dashed line). Pretty close, isn't?

### 2. Sample x Theoretical variances
```{r variance, fig.width=16}
sampleVariance <- apply(m, 1, var)
g <- ggplot(data.frame(sampleVariance), aes(sampleVariance)) +
    geom_histogram(aes(y=..density..), binwidth=1) +
    geom_density() +
    geom_vline(aes(xintercept=(1/lambda)^2, colour="Theoretical", linetype="Theoretical"), show_guide = TRUE) +
    geom_vline(aes(xintercept=var(data), colour="Simulation", linetype="Simulation"), show_guide = TRUE) +
    scale_colour_manual(name="Variances", values=c(Theoretical="blue", Simulation="red")) +
    scale_linetype_manual(name="Variances", values=c(Theoretical="dashed", Simulation="dashed"), guide=FALSE) +
    ggtitle("Sample x Theoretical variances") +
    xlab("Variance") +
    ylab("Density")
print(g)
```
The variances fits as well. The theoretical variance is (1/lambda)^2 = `r (1/lambda)^2` and our simulation indicates `r var(data)`. We have done it again!

### 3. Sample means density looks like normal distribution
```{r norma, fig.width=16}
mean <- s <- 1/lambda
error <- 1/sqrt(n)
seq <- seq(mean - 4 * s * error, mean + 4 * s * error, by=.1)
g <- ggplot(data.frame(sampleMeans), aes(sampleMeans)) +
    geom_density() +
    geom_line(data=data.frame(x=seq, y=dnorm(seq,
                                             mean=mean,
                                             sd=s*error)),
                 aes(x, y), linetype="dashed") +
    ggtitle("Sample means x Normal densities") +
    xlab("Values") +
    ylab("Density")
print(g)
```
Here is the density of our simulation (continous line) and a normal density (dashed line). The standard deviation of this normal was corrected by theoretical standard deviation * 1/sqrt(n) = `r 1/lambda * 1/sqrt(n)`.